---
title: "Final STAT 611 revisited code"
author: "Julio Carrillo"
date: "2026-06-15"
output: html_document
---

```{r}
#Import the libraries that will be used
library(dplyr)
library(ggplot2)
library(fastDummies)
library(caret)
library(tidymodels)
library(car)
library(ggfortify)
library(broom)
library(leaps)
```

```{r}
#Import the data set
df <- read.csv("Vehicle Data.csv")
```

```{r}
#Take a glimpse at the data frame to make sure it has been imported correctly
str(df)
```

```{r}
#check the column names
colnames(df)
```

```{r}
dim(df)
```

```{r}
#replace the instances of '.' and '...' with snake based column name format
names(df) <- gsub("\\.{3}","_", names(df))
names(df) <- gsub("\\.","_", names(df))
```


```{r}
colnames(df)
```

```{r}
#Denote categorical variables as factors
df$Drive_Wheels <- as.factor(df$Drive_Wheels)
df$Vehicle_Type <- as.factor(df$Vehicle_Type)
df$Number_of_Cylinders <- as.factor(df$Number_of_Cylinders)
```

```{r}
#Check if the variables have been denoted as a factor
df_factors <- df[sapply(df,is.factor)]
names(df_factors)
factor_names <- names(df_factors)
```

```{r}
#Check each level of the categorical variables
levels(df$Number_of_Cylinders)
```

```{r}
levels(df$Vehicle_Type)
```

```{r}
levels(df$Drive_Wheels)
```

```{r}
# Check the cardinality of the unique values of each categorical variable
for (v in factor_names) {
  print(table(df[[v]]))
  cat("\n")
}
```


```{r}
#Drop the values where Drive wheels or vehicle type is equal to 0
df <-filter(df, Drive_Wheels != 0 | Vehicle_Type != 0)
```


```{r}
#Check if there is any null values in the df
colSums(is.na(df))
```

```{r}
#Inspect the continuous variables in the data set
continuous_var <- names(df[sapply(df,is.numeric)])
continuous_var <- setdiff(continuous_var,"Dealer_Cost")
continuous_var
```

```{r}
#Take a glimpse of the distribution shape of continous variables
for(i in continuous_var){
  hist(df[[i]],
       probability = FALSE,
       main = paste("histogram of", i),
       xlab = i,
       col = "lightgray",
       border = "blue"
       )
}
```

```{r}
#Summary Statistics for the continuous variables
for(i in continuous_var){
  cat("summary for", i, "\n")
  print(summary(df[[i]]))
  cat("\n")
  cat("\n")
}
```

```{r}
#Distribution of the dependent variable
#Pre log transformation the distribution is skewed to the right 
ggplot(df, aes(x = MSRP)) +
  geom_histogram(bins = 30) +
  labs(title = "Distribution of MSRP", x = "MSRP", y = "Count")

#Log transformation has normalized the distribution
ggplot(df, aes(x = log(MSRP))) +
  geom_histogram(bins = 30) +
  labs(title = "Distribution of log(MSRP)", x = "log(MSRP)", y = "Count")
```
```{r}
#MSRP vs Horsepower distribtuion
ggplot(df, aes(x = Horsepower, y = MSRP)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "MSRP vs Horsepower",
    x = "Horsepower",
    y = "MSRP"
  )
```
```{r}
ggplot(df, aes(x = Vehicle_Type, y = MSRP)) +
  geom_boxplot() +
  labs(
    title = "MSRP by Vehicle Type",
    x = "Vehicle Type",
    y = "MSRP"
  )
```

```{r}
# Create weight groups
df$Weight_Group <- cut(
  df$Weight,
  breaks = quantile(df$Weight, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE),
  labels = c("Low Weight", "Medium Weight", "High Weight"),
  include.lowest = TRUE
)

# MSRP vs Horsepower by Weight group
ggplot(df, aes(x = Horsepower, y = MSRP, color = Weight_Group)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "MSRP vs Horsepower by Weight Group",
    x = "Horsepower",
    y = "MSRP",
    color = "Weight Group"
  )
```




```{r}
#Correlation amongst all the numeric predictors
numeric_df <- df %>%
  select(where(is.numeric)) %>%
  select(-Dealer_Cost)

cor(numeric_df, use = "complete.obs")
```



```{r}
#Conduct the data splitting
set.seed(1)
split <- initial_split(df,prop = 0.75, strata = MSRP)
train_df <- training(split)
test_df <- testing(split)
```

```{r}
#Check dimension of both
dim(train_df)
dim(test_df)
```

```{r}
#Calculate the percentiles and IQR range from the continous variables
bounds <- list()

for (i in continuous_var) {
  
  q1_val <- quantile(train_df[[i]], 0.25, na.rm = TRUE)
  q3_val <- quantile(train_df[[i]], 0.75, na.rm = TRUE)
  iqr_val <- IQR(train_df[[i]], na.rm = TRUE)
  median_val <- median(train_df[[i]], na.rm = TRUE)
  
  bounds[[i]] <- list(
    q1 = q1_val,
    q3 = q3_val,
    iqr = iqr_val,
    median = median_val,
    low_bounds = q1_val - 1.5 * iqr_val,
    high_bounds = q3_val + 1.5 * iqr_val
  )
}

```

```{r}
#Example of access of variables
#as.numeric is used to obtain only the number
as.numeric(bounds[["Horsepower"]]$low_bounds)
```

```{r}
#print out command to copy and paste to create the outlier indicator variable for the numeric variable
for (i in continuous_var) {
  print(
    paste0(
      "out_", i, " = as.integer(!is.na(", i, ") & (",
      i, " < as.numeric(bounds[['", i, "']]$low_bounds) | ",
      i, " > as.numeric(bounds[['", i, "']]$high_bounds)))"
    )
  )
}
```


```{r}
#Create a baseline recipe(pipeline)
recipe_baseline <- recipe(MSRP ~., data = train_df) %>%
  step_rm(Dealer_Cost, Vehicle) %>%
  
  #Outlier indicators
  step_mutate(
              out_Engine_Size = as.integer(!is.na(Engine_Size) & (Engine_Size < as.numeric(bounds[['Engine_Size']]$low_bounds) | Engine_Size > as.numeric(bounds[['Engine_Size']]$high_bounds))),
              out_Horsepower = as.integer(!is.na(Horsepower) & (Horsepower < as.numeric(bounds[['Horsepower']]$low_bounds) | Horsepower > as.numeric(bounds[['Horsepower']]$high_bounds))),
              out_Weight = as.integer(!is.na(Weight) & (Weight < as.numeric(bounds[['Weight']]$low_bounds) | Weight > as.numeric(bounds[['Weight']]$high_bounds))),
              out_Wheelbase = as.integer(!is.na(Wheelbase) & (Wheelbase < as.numeric(bounds[['Wheelbase']]$low_bounds) | Wheelbase > as.numeric(bounds[['Wheelbase']]$high_bounds))),
              out_Length = as.integer(!is.na(Length) & (Length < as.numeric(bounds[['Length']]$low_bounds) | Length > as.numeric(bounds[['Length']]$high_bounds))),
              out_Width = as.integer(!is.na(Width) & (Width < as.numeric(bounds[['Width']]$low_bounds) | Width > as.numeric(bounds[['Width']]$high_bounds)))
            ) %>%
  #missing-value indicator
  step_indicate_na(all_numeric_predictors()) %>%
  #Imputation
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  
  #Group rare categorical groups
  step_other(all_nominal_predictors(), threshold = 0.05) %>%
  #Denote the reference levels
  step_relevel(Drive_Wheels, ref_level = "FWD") %>%
  step_relevel(Vehicle_Type, ref_level = "Wagon") %>%
  step_relevel(Number_of_Cylinders, ref_level = "4") %>%
  step_mutate_at(
  c(
    Engine_Size,
    Horsepower,
    MPG_City,
    MPG_Hwy,
    Weight,
    Wheelbase,
    Length,
    Width
  ),
  fn = ~ (.x - median(.x, na.rm = TRUE)) / IQR(.x, na.rm = TRUE)
) %>%
  #OHE categorical variables
  step_dummy(all_nominal_predictors(), one_hot = FALSE) %>%
  #Remove zero variance and linear combination predictors
  step_zv(all_predictors()) %>%
  step_lincomb(all_predictors())
  
```


```{r}
#Denote the model and insert it into the workflow along with the pipeline
mod <- linear_reg() %>% set_engine("lm")
wf <-  workflow() %>% add_recipe(recipe_baseline) %>% add_model(mod)
```

```{r}
#Fit the model into the training  data
fit <- fit(wf, data=train_df)
#Extract the linear model fitted object from the workflow
lm_obj <- workflows::extract_fit_engine(fit)
```


```{r}
#prepare the k-fold cross validation
folds <- vfold_cv(train_df, v = 3)

#Evaluate the baseline on training data
res <- fit_resamples(
  wf,
  resamples = folds,
  metrics = metric_set(rmse, mae, rsq)
)

#Collect the metrics of interest 
cv_results <- collect_metrics(res) %>%
  select(.metric, mean, std_err, n) %>%
  mutate(
    mean = round(mean, 2),
    std_err = round(std_err, 2)
  )
cv_results
```

```{r}
#Outputs the main diagnostic plots for multivariate linear regression
autoplot(lm_obj)
```

```{r}
#Check the baseline vif
car::vif(lm_obj)
```



#Updated baseline 

```{r}
#Calculate quantiles and IQR for new variable which is a combination of both MPG variables to reduce multicollinearity
train_bounds <- train_df %>%
  dplyr::mutate(Combined_MPG = 0.55 * MPG_City + 0.45 * MPG_Hwy)

mpg_q1 <- quantile(train_bounds$Combined_MPG, 0.25, na.rm = TRUE)
mpg_q3 <- quantile(train_bounds$Combined_MPG, 0.75, na.rm = TRUE)
mpg_iqr <- IQR(train_bounds$Combined_MPG, na.rm = TRUE)
mpg_low <- mpg_q1 - 1.5 * mpg_iqr
mpg_high <- mpg_q3 + 1.5 * mpg_iqr
```



```{r}
#Updated recipe with the log transformation on MSRP and added MPG_combined new variable
updated_recipe <-recipe(MSRP ~., data = train_df) %>%
  step_rm(Dealer_Cost, Vehicle) %>%
  
  
  # Combine MPG variable creation and remove the both original
  step_mutate(Combined_MPG = 0.55 * MPG_City + 0.45 * MPG_Hwy) %>%
  step_rm(MPG_City, MPG_Hwy) %>%
  
  #Outlier indicators
  step_mutate(
              out_Engine_Size = as.integer(!is.na(Engine_Size) & (Engine_Size < as.numeric(bounds[['Engine_Size']]$low_bounds) | Engine_Size > as.numeric(bounds[['Engine_Size']]$high_bounds))),
              out_Horsepower = as.integer(!is.na(Horsepower) & (Horsepower < as.numeric(bounds[['Horsepower']]$low_bounds) | Horsepower > as.numeric(bounds[['Horsepower']]$high_bounds))),
              out_Weight = as.integer(!is.na(Weight) & (Weight < as.numeric(bounds[['Weight']]$low_bounds) | Weight > as.numeric(bounds[['Weight']]$high_bounds))),
              out_Wheelbase = as.integer(!is.na(Wheelbase) & (Wheelbase < as.numeric(bounds[['Wheelbase']]$low_bounds) | Wheelbase > as.numeric(bounds[['Wheelbase']]$high_bounds))),
              out_Length = as.integer(!is.na(Length) & (Length < as.numeric(bounds[['Length']]$low_bounds) | Length > as.numeric(bounds[['Length']]$high_bounds))),
              out_Width = as.integer(!is.na(Width) & (Width < as.numeric(bounds[['Width']]$low_bounds) | Width > as.numeric(bounds[['Width']]$high_bounds)))
            ) %>%
  step_mutate(
    out_Combined_MPG = as.integer(
      !is.na(Combined_MPG) & 
        (Combined_MPG < mpg_low | Combined_MPG > mpg_high)
    )
  ) %>%
  #missing-value indicator
  step_indicate_na(all_numeric_predictors()) %>%
  #Imputation
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  
  #Group rare categorical groups
  step_other(all_nominal_predictors(), threshold = 0.05) %>%
  #Denote the reference levels
  step_relevel(Drive_Wheels, ref_level = "FWD") %>%
  step_relevel(Vehicle_Type, ref_level = "Wagon") %>%
  step_relevel(Number_of_Cylinders, ref_level = "4") %>%
  step_mutate_at(c(Engine_Size, Horsepower, Weight, Wheelbase, Length, Width, Combined_MPG),
                 fn = ~ (.x - median(.x,na.rm = TRUE)) /  IQR(.x, na.rm = TRUE)) %>%
  #OHE categorical variables
  step_dummy(all_nominal_predictors(), one_hot = FALSE) %>%
  #Remove zero variance and linear combination predictors
  step_zv(all_predictors()) %>%
  step_lincomb(all_predictors()) %>%
  # Log-transform outcome during training only
  step_log(all_outcomes(), skip = TRUE)
  
```




```{r}
#update the workflow recipe
wf_updated <-  workflow() %>% add_recipe(updated_recipe) %>% add_model(mod)
```


```{r}
fit_updated <- fit(wf_updated, data=train_df)
lm_obj_updated <- workflows::extract_fit_engine(fit_updated)
```


```{r}
#Collecting metrics that are log transformed dependent variable
res_updated <- fit_resamples(
  wf_updated,
  resamples = folds,
  metrics = metric_set(rmse, mae, rsq),
  control = control_resamples(save_pred = TRUE)
)

cv_results_updated <- collect_metrics(res_updated) %>%
  select(.metric, mean, std_err, n) %>%
  mutate(
    mean = round(mean, 2),
    std_err = round(std_err, 2)
  )
cv_results_updated
```

```{r}
#transform the predictions back into dollar format mainly for MAE,and RMSE
cv_preds_updated_exp <- collect_predictions(res_updated) %>%
  mutate(.pred_dollars = exp(.pred))

cv_metrics_updated_dollars <- yardstick::metric_set(
  yardstick::rmse,
  yardstick::mae,
  yardstick::rsq
)(
  cv_preds_updated_exp,
  truth = MSRP,
  estimate = .pred_dollars
)

cv_metrics_updated_dollars
```


```{r}
autoplot(lm_obj_updated)
```




#Prepare the mallows cp on main effects selection

```{r}
#Updated recipe with main effects 
main_recipe <-recipe(MSRP ~., data = train_df) %>%
  step_rm(Dealer_Cost, Vehicle) %>%
  
  
  # Combine MPG variable creation and remove the both original
  step_mutate(Combined_MPG = 0.55 * MPG_City + 0.45 * MPG_Hwy) %>%
  step_rm(MPG_City, MPG_Hwy) %>%
  
  #Outlier indicators
  step_mutate(
              out_Engine_Size = as.integer(!is.na(Engine_Size) & (Engine_Size < as.numeric(bounds[['Engine_Size']]$low_bounds) | Engine_Size > as.numeric(bounds[['Engine_Size']]$high_bounds))),
              out_Horsepower = as.integer(!is.na(Horsepower) & (Horsepower < as.numeric(bounds[['Horsepower']]$low_bounds) | Horsepower > as.numeric(bounds[['Horsepower']]$high_bounds))),
              out_Weight = as.integer(!is.na(Weight) & (Weight < as.numeric(bounds[['Weight']]$low_bounds) | Weight > as.numeric(bounds[['Weight']]$high_bounds))),
              out_Wheelbase = as.integer(!is.na(Wheelbase) & (Wheelbase < as.numeric(bounds[['Wheelbase']]$low_bounds) | Wheelbase > as.numeric(bounds[['Wheelbase']]$high_bounds))),
              out_Length = as.integer(!is.na(Length) & (Length < as.numeric(bounds[['Length']]$low_bounds) | Length > as.numeric(bounds[['Length']]$high_bounds))),
              out_Width = as.integer(!is.na(Width) & (Width < as.numeric(bounds[['Width']]$low_bounds) | Width > as.numeric(bounds[['Width']]$high_bounds)))
            ) %>%
  step_mutate(
    out_Combined_MPG = as.integer(
      !is.na(Combined_MPG) & 
        (Combined_MPG < mpg_low | Combined_MPG > mpg_high)
    )
  ) %>%
  #missing-value indicator
  step_indicate_na(all_numeric_predictors()) %>%
  #Imputation
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  
  #Group rare categorical groups
  step_other(all_nominal_predictors(), threshold = 0.05) %>%
  #Denote the reference levels
  step_relevel(Drive_Wheels, ref_level = "FWD") %>%
  step_relevel(Vehicle_Type, ref_level = "Wagon") %>%
  step_relevel(Number_of_Cylinders, ref_level = "4") %>%
  step_mutate_at( c(Engine_Size, Horsepower, Weight, Wheelbase, Length, Width, Combined_MPG),
                 fn = ~ (.x - median(.x,na.rm = TRUE)) /  IQR(.x, na.rm = TRUE)) %>%
  #OHE categorical variables
  step_dummy(all_nominal_predictors(), one_hot = FALSE) %>%
  
  #Remove zero variance and linear combination predictors
  step_zv(all_predictors()) %>%
  step_lincomb(all_predictors()) %>%
  # Log-transform outcome during training only
  step_log(all_outcomes(), skip = TRUE)
  
```


#Prepare the data for model selection via mallows cp
```{r}
#Prep the recipe meaning we keep track of the recipe of the created pipeline
prepped_main_rec <- prep(main_recipe, training = train_df)

#bake means we apply the preprocess pipeline manually
baked_train_main <- bake(prepped_main_rec, new_data = NULL)

names(baked_train_main)
```

```{r}
# Full main effects model check that the pipeline worked
fit_lm_main_full <- lm(MSRP ~ ., data = baked_train_main)

broom::tidy(fit_lm_main_full) %>%
  dplyr::filter(is.na(estimate))

car::vif(fit_lm_main_full)
```

```{r}
#prepare the mallows cp selection with method of stepwise model selection, where we also limit the model to have up to 20 model variables
sub_main <- regsubsets(
  MSRP ~ .,
  data = baked_train_main,
  method = "seqrep",
  nbest = 1,
  nvmax = 20
)

#Have the metrics recorded within each entry or removal of a variable 
s_main <- summary(sub_main)

s_main$cp
```

```{r}
#Get the history of the adj r^2 of the model selection
s_main$adjr2
```

```{r}
#Look at the model with the lowest CP and its coefficients
id_main <- which.min(s_main$cp)
coef(sub_main, id=id_main)
```



```{r}

coefs_main <- coef(sub_main, id = id_main)
#We keep the names of the best model based on model selection into the variable excluding the intercept
vars_main <- setdiff(names(coefs_main), "(Intercept)")

#Re-fit the chosen model
form_main <- reformulate(vars_main, response = "MSRP")

#Save the lm model object into the variable
m_main <- lm(form_main, data = baked_train_main)

#tidy makes the m_int output into a table with estimate, std error, statistics, p-value
#Here we look if a variable is not estimable or problematic (high multicollinearity)
broom::tidy(m_main) %>%
  dplyr::filter(is.na(estimate))

car::vif(m_main)

```

```{r}
#re-adjust the recipe to include the best selected variables from model selection
final_main_rec <- main_recipe %>%
  #Keeps the selected variables plus the outcome variables specified
  step_select(any_of(vars_main), all_outcomes())
```


```{r}
#Re-establish the workflow
wf_main <- workflow() %>%
  add_recipe(final_main_rec) %>%
  add_model(mod)
```


```{r}
res_main <- fit_resamples(
  wf_main,
  resamples = folds,
  metrics = metric_set(rmse, mae, rsq),
  #We save the predictions in the cross validation to evaluate the trained model
  control = control_resamples(save_pred = TRUE)
)
```


```{r}
cv_preds_main <- collect_predictions(res_main) %>%
  mutate(.pred_dollars = exp(.pred))

cv_metrics_main_dollars <- yardstick::metric_set(
  yardstick::rmse,
  yardstick::mae,
  yardstick::rsq
)(
  cv_preds_main,
  truth = MSRP,
  estimate = .pred_dollars
)

cv_metrics_main_dollars
```





#Prepare meaningful interactions variables for model selection, for model selection via mallows cp

```{r}
#Updated recipe with interaction models for the 
int_recipe <-recipe(MSRP ~., data = train_df) %>%
  step_rm(Dealer_Cost, Vehicle) %>%
  
  
  # Combine MPG variable creation and remove the both original
  step_mutate(Combined_MPG = 0.55 * MPG_City + 0.45 * MPG_Hwy) %>%
  step_rm(MPG_City, MPG_Hwy) %>%
  
  #Outlier indicators
  step_mutate(
              out_Engine_Size = as.integer(!is.na(Engine_Size) & (Engine_Size < as.numeric(bounds[['Engine_Size']]$low_bounds) | Engine_Size > as.numeric(bounds[['Engine_Size']]$high_bounds))),
              out_Horsepower = as.integer(!is.na(Horsepower) & (Horsepower < as.numeric(bounds[['Horsepower']]$low_bounds) | Horsepower > as.numeric(bounds[['Horsepower']]$high_bounds))),
              out_Weight = as.integer(!is.na(Weight) & (Weight < as.numeric(bounds[['Weight']]$low_bounds) | Weight > as.numeric(bounds[['Weight']]$high_bounds))),
              out_Wheelbase = as.integer(!is.na(Wheelbase) & (Wheelbase < as.numeric(bounds[['Wheelbase']]$low_bounds) | Wheelbase > as.numeric(bounds[['Wheelbase']]$high_bounds))),
              out_Length = as.integer(!is.na(Length) & (Length < as.numeric(bounds[['Length']]$low_bounds) | Length > as.numeric(bounds[['Length']]$high_bounds))),
              out_Width = as.integer(!is.na(Width) & (Width < as.numeric(bounds[['Width']]$low_bounds) | Width > as.numeric(bounds[['Width']]$high_bounds)))
            ) %>%
  step_mutate(
    out_Combined_MPG = as.integer(
      !is.na(Combined_MPG) & 
        (Combined_MPG < mpg_low | Combined_MPG > mpg_high)
    )
  ) %>%
  #missing-value indicator
  step_indicate_na(all_numeric_predictors()) %>%
  #Imputation
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  
  #Group rare categorical groups
  step_other(all_nominal_predictors(), threshold = 0.05) %>%
  #Denote the reference levels
  step_relevel(Drive_Wheels, ref_level = "FWD") %>%
  step_relevel(Vehicle_Type, ref_level = "Wagon") %>%
  step_relevel(Number_of_Cylinders, ref_level = "4") %>%
  step_mutate_at( c(Engine_Size, Horsepower, Weight, Wheelbase, Length, Width, Combined_MPG),
                 fn = ~ (.x - median(.x,na.rm = TRUE)) /  IQR(.x, na.rm = TRUE)) %>%
  #OHE categorical variables
  step_dummy(all_nominal_predictors(), one_hot = FALSE) %>%
  #Add interactions
  step_interact(
    terms = ~ Horsepower:starts_with("Vehicle_Type_")
  ) %>%
  step_interact(
    terms = ~ Length:Width
  ) %>%
  step_interact(
    terms = ~ Horsepower:Weight
  ) %>%
  
  
  #Remove zero variance and linear combination predictors
  step_zv(all_predictors()) %>%
  step_lincomb(all_predictors()) %>%
  # Log-transform outcome during training only
  step_log(all_outcomes(), skip = TRUE)
  
```


#Prepare the data for model selection via mallows cp
```{r}
#Prep the recipe meaning we keep track of the recipe of the created pipeline
prepped_int_rec <- prep(int_recipe, training = train_df)

#bake means we apply the preprocess pipeline manually
baked_train_int <- bake(prepped_int_rec, new_data = NULL)

#Full models variable names
names(baked_train_int)
```

```{r}
# Full interaction model check that the pipeline worked
fit_lm_int_full <- lm(MSRP ~ ., data = baked_train_int)

broom::tidy(fit_lm_int_full) %>%
  dplyr::filter(is.na(estimate))

car::vif(fit_lm_int_full)
```





```{r}
#prepare the mallows cp selection with method of stepwise model selection, where we also limit the model to have up to 20 model variables
sub_int <- regsubsets(
  MSRP ~ .,
  data = baked_train_int,
  method = "seqrep",
  nbest = 1,
  nvmax = 20
)

#Have 
s_int <- summary(sub_int)

s_int$cp
```

```{r}
s_int$adjr2
```

```{r}
id_int <- which.min(s_int$cp)
coef(sub_int, id=id_int)
```

```{r}

coefs_int <- coef(sub_int, id = id_int)
#We keep the names of the best model based on model selection into the variable
vars_int <- setdiff(names(coefs_int), "(Intercept)")

#Re-fit the chosen model
form_int <- reformulate(vars_int, response = "MSRP")

#Save the lm model object into the variable
m_int <- lm(form_int, data = baked_train_int)

#tidy makes the m_int output into a table with estimate, std error, statistics, p-value
#Here we look if a variable is not estimable or problematic
broom::tidy(m_int) %>%
  dplyr::filter(is.na(estimate))

car::vif(m_int)

```



```{r}
#re-adjust the recipe 
final_int_rec <- int_recipe %>%
  step_select(any_of(vars_int), all_outcomes())
```


```{r}
#Re-establish the workflow
wf_int <- workflow() %>%
  add_recipe(final_int_rec) %>%
  add_model(mod)
```


```{r}
res_int <- fit_resamples(
  wf_int,
  resamples = folds,
  metrics = metric_set(rmse, mae, rsq),
  control = control_resamples(save_pred = TRUE)
)
```


```{r}
cv_preds_int <- collect_predictions(res_int) %>%
  mutate(.pred_dollars = exp(.pred))

cv_metrics_int_dollars <- yardstick::metric_set(
  yardstick::rmse,
  yardstick::mae,
  yardstick::rsq
)(
  cv_preds_int,
  truth = MSRP,
  estimate = .pred_dollars
)

cv_metrics_int_dollars
```


```{r}
#This has only the main effects ridge recipe
ridge_recipe <-recipe(MSRP ~., data = train_df) %>%
  step_rm(Dealer_Cost, Vehicle) %>%
  
  
  # Combine MPG variable creation and remove the both original
  step_mutate(Combined_MPG = 0.55 * MPG_City + 0.45 * MPG_Hwy) %>%
  step_rm(MPG_City, MPG_Hwy) %>%
  
  #Outlier indicators
  step_mutate(
              out_Engine_Size = as.integer(!is.na(Engine_Size) & (Engine_Size < as.numeric(bounds[['Engine_Size']]$low_bounds) | Engine_Size > as.numeric(bounds[['Engine_Size']]$high_bounds))),
              out_Horsepower = as.integer(!is.na(Horsepower) & (Horsepower < as.numeric(bounds[['Horsepower']]$low_bounds) | Horsepower > as.numeric(bounds[['Horsepower']]$high_bounds))),
              out_Weight = as.integer(!is.na(Weight) & (Weight < as.numeric(bounds[['Weight']]$low_bounds) | Weight > as.numeric(bounds[['Weight']]$high_bounds))),
              out_Wheelbase = as.integer(!is.na(Wheelbase) & (Wheelbase < as.numeric(bounds[['Wheelbase']]$low_bounds) | Wheelbase > as.numeric(bounds[['Wheelbase']]$high_bounds))),
              out_Length = as.integer(!is.na(Length) & (Length < as.numeric(bounds[['Length']]$low_bounds) | Length > as.numeric(bounds[['Length']]$high_bounds))),
              out_Width = as.integer(!is.na(Width) & (Width < as.numeric(bounds[['Width']]$low_bounds) | Width > as.numeric(bounds[['Width']]$high_bounds)))
            ) %>%
  step_mutate(
    out_Combined_MPG = as.integer(
      !is.na(Combined_MPG) & 
        (Combined_MPG < mpg_low | Combined_MPG > mpg_high)
    )
  ) %>%
  #missing-value indicator
  step_indicate_na(all_numeric_predictors()) %>%
  #Imputation
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  
  #Group rare categorical groups
  step_other(all_nominal_predictors(), threshold = 0.05) %>%
  #Denote the reference levels
  step_relevel(Drive_Wheels, ref_level = "FWD") %>%
  step_relevel(Vehicle_Type, ref_level = "Wagon") %>%
  step_relevel(Number_of_Cylinders, ref_level = "4") %>%
  step_mutate_at( c(Engine_Size, Horsepower, Weight, Wheelbase, Length, Width, Combined_MPG),
                 fn = ~ (.x - median(.x,na.rm = TRUE)) /  IQR(.x, na.rm = TRUE)) %>%
  #OHE categorical variables
  step_dummy(all_nominal_predictors(), one_hot = FALSE) %>%
  
  #Remove zero variance and linear combination predictors
  step_zv(all_predictors()) %>%
  # Log-transform outcome during training only
  step_log(all_outcomes(), skip = TRUE)
  
```


```{r}
#Ridge model specification and workflow
ridge_model <- linear_reg(penalty =  tune(), mixture = 0) %>%
  set_engine("glmnet")

wf_ridge <- workflow() %>% add_recipe(ridge_recipe) %>% add_model(ridge_model)
```


```{r}
penalty_grid <-grid_regular(penalty(), levels = 50)

res_ridge <- tune_grid(
  wf_ridge,
  resamples = folds,
  grid = penalty_grid,
  metrics = yardstick::metric_set(
    yardstick::rmse,
    yardstick::mae
  ),
  control = control_grid(save_pred = TRUE)
)
```




```{r}
# Check ridge CV results
collect_metrics(res_ridge)

# Select best penalty based on RMSE
best_ridge <- select_best(res_ridge, metric = "rmse")

best_ridge
```

```{r}
cv_preds_ridge <- collect_predictions(res_ridge) %>%
  semi_join(best_ridge, by = "penalty") %>%
  mutate(
    .pred_dollars = exp(.pred)
  )

summary(cv_preds_ridge$MSRP)
summary(cv_preds_ridge$.pred)
summary(cv_preds_ridge$.pred_dollars)
```


```{r}
cv_preds_ridge %>%
  metrics(truth = MSRP, estimate = .pred_dollars)

cv_preds_ridge %>%
  rmse(truth = MSRP, estimate = .pred_dollars)

cv_preds_ridge %>%
  mae(truth = MSRP, estimate = .pred_dollars)

cv_preds_ridge %>%
  rsq(truth = MSRP, estimate = .pred_dollars)
```



```{r}
#ridge models summary
cv_metrics_ridge_dollars <- yardstick::metric_set(
  yardstick::rmse,
  yardstick::mae,
  yardstick::rsq
)(
  cv_preds_ridge,
  truth = MSRP,
  estimate = .pred_dollars
)

cv_metrics_ridge_dollars
```



#Ridge model with interactions

```{r}
int_ridge_recipe <-recipe(MSRP ~., data = train_df) %>%
  step_rm(Dealer_Cost, Vehicle) %>%
  
  
  # Combine MPG variable creation and remove the both original
  step_mutate(Combined_MPG = 0.55 * MPG_City + 0.45 * MPG_Hwy) %>%
  step_rm(MPG_City, MPG_Hwy) %>%
  
  #Outlier indicators
  step_mutate(
              out_Engine_Size = as.integer(!is.na(Engine_Size) & (Engine_Size < as.numeric(bounds[['Engine_Size']]$low_bounds) | Engine_Size > as.numeric(bounds[['Engine_Size']]$high_bounds))),
              out_Horsepower = as.integer(!is.na(Horsepower) & (Horsepower < as.numeric(bounds[['Horsepower']]$low_bounds) | Horsepower > as.numeric(bounds[['Horsepower']]$high_bounds))),
              out_Weight = as.integer(!is.na(Weight) & (Weight < as.numeric(bounds[['Weight']]$low_bounds) | Weight > as.numeric(bounds[['Weight']]$high_bounds))),
              out_Wheelbase = as.integer(!is.na(Wheelbase) & (Wheelbase < as.numeric(bounds[['Wheelbase']]$low_bounds) | Wheelbase > as.numeric(bounds[['Wheelbase']]$high_bounds))),
              out_Length = as.integer(!is.na(Length) & (Length < as.numeric(bounds[['Length']]$low_bounds) | Length > as.numeric(bounds[['Length']]$high_bounds))),
              out_Width = as.integer(!is.na(Width) & (Width < as.numeric(bounds[['Width']]$low_bounds) | Width > as.numeric(bounds[['Width']]$high_bounds)))
            ) %>%
  step_mutate(
    out_Combined_MPG = as.integer(
      !is.na(Combined_MPG) & 
        (Combined_MPG < mpg_low | Combined_MPG > mpg_high)
    )
  ) %>%
  #missing-value indicator
  step_indicate_na(all_numeric_predictors()) %>%
  #Imputation
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  
  #Group rare categorical groups
  step_other(all_nominal_predictors(), threshold = 0.05) %>%
  #Denote the reference levels
  step_relevel(Drive_Wheels, ref_level = "FWD") %>%
  step_relevel(Vehicle_Type, ref_level = "Wagon") %>%
  step_relevel(Number_of_Cylinders, ref_level = "4") %>%
  step_mutate_at( c(Engine_Size, Horsepower, Weight, Wheelbase, Length, Width, Combined_MPG),
                 fn = ~ (.x - median(.x,na.rm = TRUE)) /  IQR(.x, na.rm = TRUE)) %>%
  #OHE categorical variables
  step_dummy(all_nominal_predictors(), one_hot = FALSE) %>%
  #Add interactions
  step_interact(
    terms = ~ Horsepower:starts_with("Vehicle_Type_")
  ) %>%
  step_interact(
    terms = ~ Length:Width
  ) %>%
  step_interact(
    terms = ~ Horsepower:Weight
  ) %>%
  
  
  #Remove zero variance and linear combination predictors
  step_zv(all_predictors()) %>%
  # Log-transform outcome during training only
  step_log(all_outcomes(), skip = TRUE)
  
```


```{r}
#Ridge model specification and workflow

wf_ridge_int <- workflow() %>% add_recipe(int_ridge_recipe) %>% add_model(ridge_model)
```


```{r}
penalty_grid <-grid_regular(penalty(), levels = 50)

res_ridge_int <- tune_grid(
  wf_ridge_int,
  resamples = folds,
  grid = penalty_grid,
  metrics = yardstick::metric_set(
    yardstick::rmse,
    yardstick::mae,
    yardstick::rsq
  ),
  control = control_grid(save_pred = TRUE)
)
```

```{r}
# Check ridge CV results
collect_metrics(res_ridge_int)

# Select best penalty based on RMSE
best_ridge_int <- select_best(res_ridge_int, metric = "rmse")

best_ridge_int
```



```{r}
cv_preds_ridge_int <- collect_predictions(res_ridge_int) %>%
  semi_join(best_ridge_int, by = "penalty") %>%
  mutate(
    .pred_dollars = exp(.pred)
  )

summary(cv_preds_ridge_int$MSRP)
summary(cv_preds_ridge_int$.pred)
summary(cv_preds_ridge_int$.pred_dollars)
```


```{r}
cv_preds_ridge_int %>%
  metrics(truth = MSRP, estimate = .pred_dollars)

cv_preds_ridge_int %>%
  rmse(truth = MSRP, estimate = .pred_dollars)

cv_preds_ridge_int %>%
  mae(truth = MSRP, estimate = .pred_dollars)

cv_preds_ridge_int %>%
  rsq(truth = MSRP, estimate = .pred_dollars)
```



```{r}

cv_metrics_ridge_int_dollars <- yardstick::metric_set(
  yardstick::rmse,
  yardstick::mae,
  yardstick::rsq
)(
  cv_preds_ridge_int,
  truth = MSRP,
  estimate = .pred_dollars
)
```



```{r}
candidate_model_summary <- bind_rows(
  cv_metrics_updated_dollars %>% mutate(Model = "Updated LM"),
  cv_metrics_main_dollars %>% mutate(Model = "Main Effects LM"),
  cv_metrics_int_dollars %>% mutate(Model = "Interaction LM"),
  cv_metrics_ridge_dollars %>% mutate(Model = "Ridge Main Effects"),
  cv_metrics_ridge_int_dollars %>% mutate(Model = "Ridge Interaction")
) %>%
  select(Model, .metric, .estimate) %>%
  mutate(
    .metric = case_when(
      .metric == "rmse" ~ "RMSE",
      .metric == "mae"  ~ "MAE",
      .metric == "rsq"  ~ "R-squared"
    ),
    .estimate = round(.estimate, 2)
  ) %>%
  pivot_wider(
    names_from = .metric,
    values_from = .estimate
  ) %>%
  arrange(RMSE)

candidate_model_summary
```



#Evaluation of the final model based on the evaluation metrics
```{r}
#Important: do not prep the recipe again on the test data.
# Bake training and testing data
#Training does the model fit and the hypothesis testing
baked_train_int <- bake(prepped_int_rec, new_data = train_df)
baked_test_int  <- bake(prepped_int_rec, new_data = test_df)
```



```{r}
#Based on the evaluated models, the interaction linear models
#But we need to force the interactions
# Manually force important interaction terms
forced_interactions <- c(
  "Length",

  "Vehicle_Type_Minivan",
  "Vehicle_Type_Pickup",
  "Vehicle_Type_Sedan",
  "Vehicle_Type_Sports.Car",
  "Vehicle_Type_SUV",

  "Horsepower_x_Vehicle_Type_Minivan",
  "Horsepower_x_Vehicle_Type_Pickup",
  "Horsepower_x_Vehicle_Type_Sedan",
  "Horsepower_x_Vehicle_Type_Sports.Car",
  "Horsepower_x_Vehicle_Type_SUV"
)

# Final variable set
vars_final <- union(vars_int, forced_interactions)

# Check missing columns
setdiff(vars_final, names(baked_train_int))
```




```{r}
# Create final formula
form_final <- reformulate(vars_final, response = "MSRP")

# Fit final model
final_model <- lm(form_final, data = baked_train_int)

# Hypothesis tests / coefficient table
summary(final_model)
```


```{r}
vif(final_model)
```

```{r}
# Predict on test set
test_preds_final <- predict(final_model, newdata = baked_test_int)

# Test metrics
yardstick::metrics(
  data.frame(
    truth = baked_test_int$MSRP,
    estimate = test_preds_final
  ),
  truth = truth,
  estimate = estimate
)
```

```{r}
#Corrected
# Create final formula with log outcome
form_final <- as.formula(
  paste("log(MSRP) ~", paste(vars_final, collapse = " + "))
)

# Fit final model
final_model <- lm(form_final, data = baked_train_int)

# Hypothesis tests / coefficient table
summary(final_model)
```

```{r}
vif(final_model)
```

```{r}
autoplot(final_model)
```

#Length only added final model


```{r}
#Based on the evaluated models, the interaction linear models
#But we need to force the interactions
# Manually force important interaction terms
Additional_variables<- c(
  "Length"
)

# Final variable set
vars_final <- union(vars_int, Additional_variables)

# Check missing columns
setdiff(vars_final, names(baked_train_int))
```




```{r}
# Create final formula
form_final <- reformulate(vars_final, response = "MSRP")

# Fit final model
final_model <- lm(form_final, data = baked_train_int)

# Hypothesis tests / coefficient table
summary(final_model)
```

```{r}
form_final_log <- as.formula(
  paste("log(MSRP) ~", paste(vars_final, collapse = " + "))
)

final_model_log <- lm(form_final_log, data = baked_train_int)

summary(final_model_log)
autoplot(final_model_log)
```


```{r}
autoplot(final_model_log)
```


```{r}
vif(final_model_log)
```





```{r}
# Predict on log scale
test_preds_log <- predict(final_model_log, newdata = baked_test_int)

# Test metrics on log scale
yardstick::metrics(
  data.frame(
    truth = log(baked_test_int$MSRP),
    estimate = test_preds_log
  ),
  truth = truth,
  estimate = estimate
)
```



```{r}
# Predict on log scale
test_preds_log <- predict(final_model_log, newdata = baked_test_int)

# Convert back to dollar scale
test_preds_dollar <- exp(test_preds_log)

# Test metrics on dollar scale
yardstick::metrics(
  data.frame(
    truth = baked_test_int$MSRP,
    estimate = test_preds_dollar
  ),
  truth = truth,
  estimate = estimate
)
```
